---
title: "Bayesian Generalized Linear Mixed Models for Trajectory Prevalences"
date: "`r Sys.Date()`"
encoding: UTF-8
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    toc_depth: 4
    number_sections: true
    code_folding: hide
  word_document:
    toc: true
    toc_depth: '4'
  pdf_document:
    toc: true
    toc_depth: '4'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = FALSE)

# Set the root directory for the entire document (recommended way)
knitr::opts_knit$set(root.dir = dirname(rstudioapi::getActiveDocumentContext()$path))
```

# Preparation

```{r load-packages}
library(readxl)
library(readr)
library(dplyr)
library(gt)
library(purrr)
library(tidyr)
library(brms)
library(loo)
library(posterior)
library(bayesplot)
```

## Load Data

```{r load-data}
# load data
df_moderation <- read_excel("../extracted_data/Clinical_extraction_final.xlsx")

# force all character columns to UTF-8
df_moderation <- df_moderation |>
  dplyr::mutate(dplyr::across(where(is.character), ~ enc2utf8(.x)))

# select relevant columns
df_moderation <- df_moderation[, c("Study", "Sample_name", "MID", "Sample_size", "Mean_age", "SD", "Percentage_women", "trauma_type", "TP_assessments", "Treatment_type", "Treatment_length", "N_trajectories", "Persistent", "Persistent_n", "Persistent_percentage", "Partially_Decreasing", "Partially_Decreasing_n", "Partially_Decreasing_percentage", "Decreasing", "Decreasing_n", "Decreasing_percentage", "Relapsing", "Relapsing_n", "Relapsing_percentage")] 

# Display the first rows to check the data
head(df_moderation)
```

## Descriptive Statistics

We provide basic descriptive statistics of the included studies, without excluding any data due to missingness.

```{r descriptives}
# Calculate descriptives
total_number_of_samples <- nrow(df_moderation)
total_Sample_size <- sum(df_moderation$Sample_size, na.rm = TRUE)

# Create a summary table
summary_table <- data.frame(
  Description = c("Number of studies / cohorts", "Total sample size (sum of available Sample_size)"),
  Value = c(
    format(total_number_of_samples, big.mark = ",", trim = TRUE), 
    format(total_Sample_size, big.mark = ",", trim = TRUE)
  )
)

# Display as a table
knitr::kable(summary_table, caption = "Descriptive Statistics of the Dataset")
```

### Descriptives Tables for All Studies

```{r descriptives-studies-table}
overview <- tibble(
  `Author(s), Year`                = df_moderation$Study,
  `Sample Name`                    = df_moderation$Sample_name,
  `Sample Size`                    = df_moderation$Sample_size,
  `Mean Age`                       = df_moderation$Mean_age,
  `Percentage Women`               = df_moderation$Percentage_women,
  `Assessed Trauma Type`           = df_moderation$trauma_type,
  `Number of TP assessments`       = df_moderation$TP_assessments,
  `Treatment Type`                 = df_moderation$Treatment_type,
  `Treatment Length (weeks)`       = df_moderation$Treatment_length,
  `Number of Trajectories Found`   = df_moderation$N_trajectories
) |>
  arrange(`Author(s), Year`)

if (knitr::is_html_output() || knitr::is_latex_output()) {

  fmt <- if (knitr::is_latex_output()) "latex" else "html"

  overview |>
    mutate(
      `Sample Size` = format(as.numeric(`Sample Size`),
                             big.mark = ",", trim = TRUE),

      MeanAge_num = parse_number(as.character(`Mean Age`)),
      `Mean Age` = if_else(
        is.na(MeanAge_num),
        NA_character_,
        sprintf("%.1f", MeanAge_num)
      ),

      Women_num = parse_number(as.character(`Percentage Women`)),
      `Percentage Women` = if_else(
        is.na(Women_num),
        NA_character_,
        sprintf("%.1f%%", Women_num)
      )
    ) |>
    select(-MeanAge_num, -Women_num) |>
    knitr::kable(
      format   = fmt,
      caption  = "General Descriptive Variables per Sample",
      booktabs = TRUE,
      linesep  = ""
    ) |>
    kableExtra::kable_styling(
      full_width = FALSE,
      position   = "left",
      bootstrap_options = c("striped", "hover", "condensed")
    ) |>
    kableExtra::column_spec(1, bold = TRUE) |>
    kableExtra::row_spec(
      0,
      extra_css = "font-family: 'Times New Roman', Times, serif;"
    ) |>
    kableExtra::row_spec(
      1:nrow(overview),
      extra_css = "font-family: 'Times New Roman', Times, serif;"
    ) |>
    kableExtra::scroll_box(height = "470px")
} else {
  overview
}
```

# Prevalences

## Bayesian GLMM Function

We fit a Bayesian random-intercept GLMM using `brms`. This models the binomial data-generating process via Hamiltonian Monte Carlo (Stan). This provides fully probabilistic uncertainty intervals (credible intervals) and behaves gracefully when the number of studies is small, between-study heterogeneity is close to zero, or studies report extreme proportions.

We use weakly informative priors: `normal(0, 1.5)` on the population-level logit-scale intercept and a `student_t(3, 0, 1)` prior on the between-study standard deviation. The function returns a list containing both the summary table data frame and the full `brmsfit` object (used later for Leave-One-Out Cross-Validation).

```{r portable-bglmm-function}
# Portable function for pooled prevalence via Bayesian GLMM (brms)
fit_bglmm_prev <- function(df, event_col, n_col = "Sample_size", label = event_col,
                           chains = 4, iter = 4000, warmup = 1000,
                           adapt_delta = 0.95, seed = 123, refresh = 0, min_k = 3) {

  xi <- suppressWarnings(round(as.numeric(df[[event_col]])))
  ni <- suppressWarnings(as.numeric(df[[n_col]]))

  ok <- !is.na(xi) & !is.na(ni) & ni > 0 & xi >= 0 & xi <= ni
  k  <- length(unique(df$Study[ok]))

  if (sum(ok) < min_k) {
    message(label, ": only ", sum(ok),
            " usable studies after filtering (minimum required: ", min_k,
            "); Bayesian GLMM not estimable. Skipping.")
    return(
      list(
        summary = data.frame(
          Analysis = label, Pooled_Prevalence = NA_real_, CrI_Lower = NA_real_,
          CrI_Upper = NA_real_, Samples_included = sum(ok), Logit_Estimate = NA_real_,
          Tau = NA_real_, Rhat_max = NA_real_, check.names = FALSE
        ),
        model = NULL
      )
    )
  }

  df_use <- df[ok, , drop = FALSE]
  df_use$n_events <- xi[ok]
  df_use$n_total  <- ni[ok]

  priors <- c(
    set_prior("normal(0, 1.5)", class = "Intercept"),
    set_prior("student_t(3, 0, 1)", class = "sd")
  )

  fit <- brm(
    n_events | trials(n_total) ~ 1 + (1 | Study),
    data    = df_use,
    family  = binomial(link = "logit"),
    prior   = priors,
    chains  = chains,
    iter    = iter,
    warmup  = warmup,
    seed    = seed,
    refresh = refresh,
    control = list(adapt_delta = adapt_delta)
  )

  draws      <- as_draws_df(fit)
  post_int   <- draws$b_Intercept
  prev_draws <- plogis(post_int)
  tau_draws  <- draws$sd_Study__Intercept

  summary_df <- data.frame(
    Analysis          = label,
    Pooled_Prevalence = round(mean(prev_draws), 3),
    CrI_Lower         = round(as.numeric(quantile(prev_draws, .025)), 3),
    CrI_Upper         = round(as.numeric(quantile(prev_draws, .975)), 3),
    Samples_included  = k,
    Logit_Estimate    = round(mean(post_int), 3),
    Tau               = round(mean(tau_draws), 3),
    Rhat_max          = round(max(brms::rhat(fit), na.rm = TRUE), 3),
    check.names = FALSE
  )
  
  return(list(summary = summary_df, model = fit))
}
```

## Conservative (Zero-Imputed) Estimate Helper

The main analysis above only includes studies that identified a given trajectory as a distinct latent class. This means the pooled prevalence for less-commonly-detected trajectories (e.g. Decreasing, Partially Decreasing) is estimated only among the subset of studies capable of finding that pattern — which is not the same as the prevalence across the full body of evidence, and may overestimate the "true" prevalence.

As a more conservative complement, we build a second dataset per trajectory in which **every study with a usable total sample size is retained**, and studies that did not identify the trajectory in question are coded as contributing **0 events** (rather than being dropped). Pooling across this dataset gives a lower-bound estimate; the main analysis serves as the corresponding upper-bound estimate.

```{r conservative-helper}
# Build a "conservative" (zero-imputed) version of a trajectory dataset.
# Studies with a valid Sample_size but no identified trajectory (NA percentage)
# are retained with n_events = 0, instead of being excluded.
make_conservative_df <- function(df, event_col, percentage_col, n_col = "Sample_size") {
  df_cons <- df
  has_size <- !is.na(df_cons[[n_col]]) & df_cons[[n_col]] > 0
  df_cons  <- df_cons[has_size, , drop = FALSE]

  df_cons[[event_col]] <- ifelse(
    is.na(df_cons[[percentage_col]]),
    0,
    df_cons[[event_col]]
  )

  df_cons
}
```

---

## Decreasing Symptoms Trajectory

### Descriptives

```{r Decreasing-descriptives}
df_Decreasing <- df_moderation
df_Decreasing$Decreasing_percentage <- as.numeric(df_Decreasing$Decreasing_percentage)
df_Decreasing$Decreasing_n <- round((df_Decreasing$Decreasing_percentage / 100) * df_Decreasing$Sample_size)

total_Decreasing_n <- sum(df_Decreasing$Decreasing_n, na.rm = TRUE)
unique_Decreasing_studies <- length(unique(df_Decreasing$Study[!is.na(df_Decreasing$Decreasing_percentage)]))

additional_info <- data.frame(
  Description = c("Total number of individuals in Decreasing trajectory", "Number of unique studies"),
  Value = format(c(total_Decreasing_n, unique_Decreasing_studies), big.mark = ",")
)
knitr::kable(additional_info, caption = "Additional Descriptives for Decreasing Trajectory Analysis")
```

### Main Analysis

```{r Decreasing-main, results='hide'}
res_Decreasing <- fit_bglmm_prev(df_Decreasing, event_col = "Decreasing_n", 
                                 n_col = "Sample_size", label = "Main Analysis")
bglmm_Decreasing <- res_Decreasing$summary
```
```{r Decreasing-main-table}
knitr::kable(bglmm_Decreasing, caption = "Pooled Prevalence Estimate for Decreasing Trajectory")
```

### Sensitivity Analysis 1 (Exclude ≥90%)

```{r Decreasing-sens1, results='hide'}
df_Decreasing_sens1 <- subset(df_Decreasing, !is.na(Decreasing_n) & !is.na(Sample_size) &
                              !is.na(Decreasing_percentage) & Decreasing_percentage < 90)

bglmm_Decreasing_sens1 <- fit_bglmm_prev(df_Decreasing_sens1, "Decreasing_n", "Sample_size", 
                                         "Sensitivity 1: Exclude ≥90% Decreasing", min_k = 5)$summary
```
```{r Decreasing-sens1-table}
knitr::kable(bglmm_Decreasing_sens1, caption = "Sensitivity 1: Excl. ≥90% Prevalence")
```

### Sensitivity Analysis 2 (Large Samples, N > 999)

```{r Decreasing-sens2, results='hide'}
df_Decreasing_sens2 <- subset(df_Decreasing, !is.na(Decreasing_n) & !is.na(Sample_size) &
                              Sample_size > 999)

bglmm_Decreasing_sens2 <- fit_bglmm_prev(df_Decreasing_sens2, "Decreasing_n", "Sample_size", 
                                         "Sensitivity 2: N > 999", min_k = 5)$summary
```
```{r Decreasing-sens2-table}
knitr::kable(bglmm_Decreasing_sens2, caption = "Sensitivity 2: Samples > 999")
```

### Sensitivity Analysis 3 (Exclude Influential Studies via Pareto *k*)

We use Bayesian Leave-One-Out (LOO) cross-validation to identify highly influential studies. Studies with a Pareto *k* shape parameter > 0.7 are considered highly influential and are excluded in this sensitivity step.

```{r Decreasing-influential, results='hide'}
# Run LOO on the main model fit
loo_Decreasing <- brms::loo(res_Decreasing$model)
pareto_k_Decreasing <- loo_Decreasing$diagnostics$pareto_k

# Identify studies with pareto k > 0.7
influential_indices <- which(pareto_k_Decreasing > 0.7)
influential_studies_Decreasing <- unique(res_Decreasing$model$data$Study[influential_indices])

# Exclude and refit
df_Decreasing_noinf <- df_Decreasing %>%
  filter(!Study %in% influential_studies_Decreasing)

bglmm_Decreasing_noinf <- fit_bglmm_prev(df_Decreasing_noinf, "Decreasing_n", "Sample_size",
                           "Sensitivity 3: Excluding Influential Studies", min_k = 5)$summary
```
```{r Decreasing-influential-table}
knitr::kable(data.frame(Study = influential_studies_Decreasing), caption = "Influential Studies Excluded (Pareto k > 0.7)")
knitr::kable(bglmm_Decreasing_noinf, caption = "Sensitivity 3: Excl. Influential Studies")
```

### Sensitivity Analysis 4 (Conservative: Zero-Imputed Lower Bound)

```{r Decreasing-conservative, results='hide'}
df_Decreasing_cons <- make_conservative_df(df_Decreasing, "Decreasing_n", "Decreasing_percentage")

bglmm_Decreasing_cons <- fit_bglmm_prev(df_Decreasing_cons, "Decreasing_n", "Sample_size",
                           "Sensitivity 4: Conservative (Zero-Imputed)", min_k = 5)$summary
```
```{r Decreasing-conservative-table}
knitr::kable(bglmm_Decreasing_cons, caption = "Sensitivity 4: Conservative Lower-Bound Estimate (Zero-Imputed)")
```

---

## Partially Decreasing Symptom Trajectory

### Descriptives

```{r Partially_Decreasing-descriptives}
df_Partially_Decreasing <- df_moderation
df_Partially_Decreasing$Partially_Decreasing_percentage <- as.numeric(df_Partially_Decreasing$Partially_Decreasing_percentage)
df_Partially_Decreasing$Partially_Decreasing_n <- round((df_Partially_Decreasing$Partially_Decreasing_percentage / 100) * df_Partially_Decreasing$Sample_size)

total_Partially_Decreasing_n <- sum(df_Partially_Decreasing$Partially_Decreasing_n, na.rm = TRUE)
unique_Partially_Decreasing_studies <- length(unique(df_Partially_Decreasing$Study[!is.na(df_Partially_Decreasing$Partially_Decreasing_percentage)]))

additional_info <- data.frame(
  Description = c("Total number of individuals in Partially Decreasing trajectory", "Number of unique studies"),
  Value = format(c(total_Partially_Decreasing_n, unique_Partially_Decreasing_studies), big.mark = ",")
)
knitr::kable(additional_info, caption = "Additional Descriptives for Partially Decreasing Trajectory Analysis")
```

### Main Analysis

```{r Partially_Decreasing-main, results='hide'}
res_Partially_Decreasing <- fit_bglmm_prev(df_Partially_Decreasing, event_col = "Partially_Decreasing_n", 
                                           n_col = "Sample_size", label = "Main Analysis")
bglmm_Partially_Decreasing <- res_Partially_Decreasing$summary
```
```{r Partially_Decreasing-main-table}
knitr::kable(bglmm_Partially_Decreasing, caption = "Pooled Prevalence Estimate for Partially Decreasing Trajectory")
```

### Sensitivity Analysis 1 (Exclude ≥90%)

```{r Partially_Decreasing-sens1, results='hide'}
df_Partially_Decreasing_sens1 <- subset(df_Partially_Decreasing, !is.na(Partially_Decreasing_n) & !is.na(Sample_size) &
                              !is.na(Partially_Decreasing_percentage) & Partially_Decreasing_percentage < 90)

bglmm_Partially_Decreasing_sens1 <- fit_bglmm_prev(df_Partially_Decreasing_sens1, "Partially_Decreasing_n", "Sample_size", 
                                         "Sensitivity 1: Exclude ≥90% Partially Decreasing", min_k = 5)$summary
```
```{r Partially_Decreasing-sens1-table}
knitr::kable(bglmm_Partially_Decreasing_sens1, caption = "Sensitivity 1: Excl. ≥90% Prevalence")
```

### Sensitivity Analysis 2 (Large Samples, N > 999)

```{r Partially_Decreasing-sens2, results='hide'}
df_Partially_Decreasing_sens2 <- subset(df_Partially_Decreasing, !is.na(Partially_Decreasing_n) & !is.na(Sample_size) &
                              Sample_size > 999)

bglmm_Partially_Decreasing_sens2 <- fit_bglmm_prev(df_Partially_Decreasing_sens2, "Partially_Decreasing_n", "Sample_size", 
                                         "Sensitivity 2: N > 999", min_k = 5)$summary
```
```{r Partially_Decreasing-sens2-table}
knitr::kable(bglmm_Partially_Decreasing_sens2, caption = "Sensitivity 2: Samples > 999")
```

### Sensitivity Analysis 3 (Exclude Influential Studies via Pareto *k*)

```{r Partially_Decreasing-influential, results='hide'}
loo_Partially_Decreasing <- brms::loo(res_Partially_Decreasing$model)
pareto_k_Partially <- loo_Partially_Decreasing$diagnostics$pareto_k

influential_indices <- which(pareto_k_Partially > 0.7)
influential_studies_Partially_Decreasing <- unique(res_Partially_Decreasing$model$data$Study[influential_indices])

df_Partially_Decreasing_noinf <- df_Partially_Decreasing %>%
  filter(!Study %in% influential_studies_Partially_Decreasing)

bglmm_Partially_Decreasing_noinf <- fit_bglmm_prev(df_Partially_Decreasing_noinf, "Partially_Decreasing_n", "Sample_size",
                           "Sensitivity 3: Excluding Influential Studies", min_k = 5)$summary
```
```{r Partially_Decreasing-influential-table}
knitr::kable(data.frame(Study = influential_studies_Partially_Decreasing), caption = "Influential Studies Excluded (Pareto k > 0.7)")
knitr::kable(bglmm_Partially_Decreasing_noinf, caption = "Sensitivity 3: Excl. Influential Studies")
```

### Sensitivity Analysis 4 (Conservative: Zero-Imputed Lower Bound)

```{r Partially_Decreasing-conservative, results='hide'}
df_Partially_Decreasing_cons <- make_conservative_df(df_Partially_Decreasing, "Partially_Decreasing_n", "Partially_Decreasing_percentage")

bglmm_Partially_Decreasing_cons <- fit_bglmm_prev(df_Partially_Decreasing_cons, "Partially_Decreasing_n", "Sample_size",
                           "Sensitivity 4: Conservative (Zero-Imputed)", min_k = 5)$summary
```
```{r Partially_Decreasing-conservative-table}
knitr::kable(bglmm_Partially_Decreasing_cons, caption = "Sensitivity 4: Conservative Lower-Bound Estimate (Zero-Imputed)")
```

---

## Persistent Symptoms Trajectory

### Descriptives

```{r Persistent-descriptives}
df_Persistent <- df_moderation
df_Persistent$Persistent_percentage <- as.numeric(df_Persistent$Persistent_percentage)
df_Persistent$Persistent_n <- round((df_Persistent$Persistent_percentage / 100) * df_Persistent$Sample_size)

total_Persistent_n <- sum(df_Persistent$Persistent_n, na.rm = TRUE)
unique_Persistent_studies <- length(unique(df_Persistent$Study[!is.na(df_Persistent$Persistent_percentage)]))

additional_info <- data.frame(
  Description = c("Total number of individuals in Persistent trajectory", "Number of unique studies"),
  Value = format(c(total_Persistent_n, unique_Persistent_studies), big.mark = ",")
)
knitr::kable(additional_info, caption = "Additional Descriptives for Persistent Trajectory Analysis")
```

### Main Analysis

```{r Persistent-main, results='hide'}
res_Persistent <- fit_bglmm_prev(df_Persistent, event_col = "Persistent_n", 
                                 n_col = "Sample_size", label = "Main Analysis")
bglmm_Persistent <- res_Persistent$summary
```
```{r Persistent-main-table}
knitr::kable(bglmm_Persistent, caption = "Pooled Prevalence Estimate for Persistent Trajectory")
```

### Sensitivity Analysis 1 (Exclude ≥90%)

```{r Persistent-sens1, results='hide'}
df_Persistent_sens1 <- subset(df_Persistent, !is.na(Persistent_n) & !is.na(Sample_size) &
                              !is.na(Persistent_percentage) & Persistent_percentage < 90)

bglmm_Persistent_sens1 <- fit_bglmm_prev(df_Persistent_sens1, "Persistent_n", "Sample_size", 
                                         "Sensitivity 1: Exclude ≥90% Persistent", min_k = 5)$summary
```
```{r Persistent-sens1-table}
knitr::kable(bglmm_Persistent_sens1, caption = "Sensitivity 1: Excl. ≥90% Prevalence")
```

### Sensitivity Analysis 2 (Large Samples, N > 999)

```{r Persistent-sens2, results='hide'}
df_Persistent_sens2 <- subset(df_Persistent, !is.na(Persistent_n) & !is.na(Sample_size) &
                              Sample_size > 999)

bglmm_Persistent_sens2 <- fit_bglmm_prev(df_Persistent_sens2, "Persistent_n", "Sample_size", 
                                         "Sensitivity 2: N > 999", min_k = 5)$summary
```
```{r Persistent-sens2-table}
knitr::kable(bglmm_Persistent_sens2, caption = "Sensitivity 2: Samples > 999")
```

### Sensitivity Analysis 3 (Exclude Influential Studies via Pareto *k*)

```{r Persistent-influential, results='hide'}
loo_Persistent <- brms::loo(res_Persistent$model)
pareto_k_Persistent <- loo_Persistent$diagnostics$pareto_k

influential_indices <- which(pareto_k_Persistent > 0.7)
influential_studies_Persistent <- unique(res_Persistent$model$data$Study[influential_indices])

df_Persistent_noinf <- df_Persistent %>%
  filter(!Study %in% influential_studies_Persistent)

bglmm_Persistent_noinf <- fit_bglmm_prev(df_Persistent_noinf, "Persistent_n", "Sample_size",
                           "Sensitivity 3: Excluding Influential Studies", min_k = 5)$summary
```
```{r Persistent-influential-table}
knitr::kable(data.frame(Study = influential_studies_Persistent), caption = "Influential Studies Excluded (Pareto k > 0.7)")
knitr::kable(bglmm_Persistent_noinf, caption = "Sensitivity 3: Excl. Influential Studies")
```

### Sensitivity Analysis 4 (Conservative: Zero-Imputed Lower Bound)

```{r Persistent-conservative, results='hide'}
df_Persistent_cons <- make_conservative_df(df_Persistent, "Persistent_n", "Persistent_percentage")

bglmm_Persistent_cons <- fit_bglmm_prev(df_Persistent_cons, "Persistent_n", "Sample_size",
                           "Sensitivity 4: Conservative (Zero-Imputed)", min_k = 5)$summary
```
```{r Persistent-conservative-table}
knitr::kable(bglmm_Persistent_cons, caption = "Sensitivity 4: Conservative Lower-Bound Estimate (Zero-Imputed)")
```

---

## Relapsing Symptoms Trajectory

### Descriptives

```{r Relapsing-descriptives}
df_Relapsing <- df_moderation
df_Relapsing$Relapsing_percentage <- as.numeric(df_Relapsing$Relapsing_percentage)
df_Relapsing$Relapsing_n <- round((df_Relapsing$Relapsing_percentage / 100) * df_Relapsing$Sample_size)

total_Relapsing_n <- sum(df_Relapsing$Relapsing_n, na.rm = TRUE)
unique_Relapsing_studies <- length(unique(df_Relapsing$Study[!is.na(df_Relapsing$Relapsing_percentage)]))

additional_info <- data.frame(
  Description = c("Total number of individuals in Relapsing trajectory", "Number of unique studies"),
  Value = format(c(total_Relapsing_n, unique_Relapsing_studies), big.mark = ",")
)
knitr::kable(additional_info, caption = "Additional Descriptives for Relapsing Trajectory Analysis")
```

### Main Analysis

```{r Relapsing-main, results='hide'}
res_Relapsing <- fit_bglmm_prev(df_Relapsing, event_col = "Relapsing_n", 
                                n_col = "Sample_size", label = "Main Analysis")
bglmm_Relapsing <- res_Relapsing$summary
```
```{r Relapsing-main-table}
knitr::kable(bglmm_Relapsing, caption = "Pooled Prevalence Estimate for Relapsing Trajectory")
```

### Sensitivity Analysis 1 (Exclude ≥90%)

```{r Relapsing-sens1, results='hide'}
df_Relapsing_sens1 <- subset(df_Relapsing, !is.na(Relapsing_n) & !is.na(Sample_size) &
                              !is.na(Relapsing_percentage) & Relapsing_percentage < 90)

bglmm_Relapsing_sens1 <- fit_bglmm_prev(df_Relapsing_sens1, "Relapsing_n", "Sample_size", 
                                         "Sensitivity 1: Exclude ≥90% Relapsing", min_k = 5)$summary
```
```{r Relapsing-sens1-table}
knitr::kable(bglmm_Relapsing_sens1, caption = "Sensitivity 1: Excl. ≥90% Prevalence")
```

### Sensitivity Analysis 2 (Large Samples, N > 999)

```{r Relapsing-sens2, results='hide'}
df_Relapsing_sens2 <- subset(df_Relapsing, !is.na(Relapsing_n) & !is.na(Sample_size) &
                              Sample_size > 999)

bglmm_Relapsing_sens2 <- fit_bglmm_prev(df_Relapsing_sens2, "Relapsing_n", "Sample_size", 
                                         "Sensitivity 2: N > 999", min_k = 5)$summary
```
```{r Relapsing-sens2-table}
knitr::kable(bglmm_Relapsing_sens2, caption = "Sensitivity 2: Samples > 999")
```

### Sensitivity Analysis 3 (Exclude Influential Studies via Pareto *k*)

```{r Relapsing-influential, results='hide'}
loo_Relapsing <- brms::loo(res_Relapsing$model)
pareto_k_Relapsing <- loo_Relapsing$diagnostics$pareto_k

influential_indices <- which(pareto_k_Relapsing > 0.7)
influential_studies_Relapsing <- unique(res_Relapsing$model$data$Study[influential_indices])

df_Relapsing_noinf <- df_Relapsing %>%
  filter(!Study %in% influential_studies_Relapsing)

bglmm_Relapsing_noinf <- fit_bglmm_prev(df_Relapsing_noinf, "Relapsing_n", "Sample_size",
                           "Sensitivity 3: Excluding Influential Studies", min_k = 5)$summary
```
```{r Relapsing-influential-table}
knitr::kable(data.frame(Study = influential_studies_Relapsing), caption = "Influential Studies Excluded (Pareto k > 0.7)")
knitr::kable(bglmm_Relapsing_noinf, caption = "Sensitivity 3: Excl. Influential Studies")
```

### Sensitivity Analysis 4 (Conservative: Zero-Imputed Lower Bound)

```{r Relapsing-conservative, results='hide'}
df_Relapsing_cons <- make_conservative_df(df_Relapsing, "Relapsing_n", "Relapsing_percentage")

bglmm_Relapsing_cons <- fit_bglmm_prev(df_Relapsing_cons, "Relapsing_n", "Sample_size",
                           "Sensitivity 4: Conservative (Zero-Imputed)", min_k = 5)$summary
```
```{r Relapsing-conservative-table}
knitr::kable(bglmm_Relapsing_cons, caption = "Sensitivity 4: Conservative Lower-Bound Estimate (Zero-Imputed)")
```

---

# Tables for Paper

## Prevalences Summary Table

We now present the pooled prevalence per trajectory alongside its conservative (zero-imputed) counterpart, giving an upper- and lower-bound estimate for each trajectory.

```{r prevalences-summary-table, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
# Combine already computed Bayesian main results to save time
fits_df <- bind_rows(
  bglmm_Decreasing |> mutate(trajectory = "Decreasing symptoms"),
  bglmm_Partially_Decreasing |> mutate(trajectory = "Partially decreasing symptoms"),
  bglmm_Persistent |> mutate(trajectory = "Persistent symptoms"),
  bglmm_Relapsing |> mutate(trajectory = "Relapsing symptoms")
)

# Conservative (zero-imputed) lower-bound estimates
fits_cons_df <- bind_rows(
  bglmm_Decreasing_cons |> mutate(trajectory = "Decreasing symptoms"),
  bglmm_Partially_Decreasing_cons |> mutate(trajectory = "Partially decreasing symptoms"),
  bglmm_Persistent_cons |> mutate(trajectory = "Persistent symptoms"),
  bglmm_Relapsing_cons |> mutate(trajectory = "Relapsing symptoms")
) |>
  transmute(
    trajectory,
    cons_prev = sprintf("%.1f%%", 100 * Pooled_Prevalence),
    cons_cri  = sprintf("%.1f to %.1f%%", 100 * CrI_Lower, 100 * CrI_Upper)
  )

# Extract totals dynamically based on the dataset logic for N(%)
calc_n <- function(df, perc_col) {
  event_col <- round((as.numeric(df[[perc_col]]) / 100) * df[["Sample_size"]])
  sum(event_col, na.rm = TRUE)
}

totals <- tibble(
  trajectory = c("Decreasing symptoms", "Partially decreasing symptoms", "Persistent symptoms", "Relapsing symptoms"),
  N = c(
    calc_n(df_moderation, "Decreasing_percentage"),
    calc_n(df_moderation, "Partially_Decreasing_percentage"),
    calc_n(df_moderation, "Persistent_percentage"),
    calc_n(df_moderation, "Relapsing_percentage")
  )
) %>%
  mutate(pct = round(100 * N / sum(N, na.rm = TRUE), 1))

# Format the final table frame
table1_df <- fits_df %>%
  left_join(totals, by = "trajectory") %>%
  left_join(fits_cons_df, by = "trajectory") %>%
  transmute(
    trajectory,
    k = as.integer(Samples_included),
    N_pct = sprintf("%s (%.1f%%)", format(N, big.mark = ","), pct),
    bayes_prev = sprintf("%.1f%%", 100 * Pooled_Prevalence),
    bayes_cri  = sprintf("%.1f to %.1f%%", 100 * CrI_Lower, 100 * CrI_Upper),
    cons_prev,
    cons_cri,
    logit = round(Logit_Estimate, 2),
    Tau = round(Tau, 2),
    Rhat_max = round(Rhat_max, 3)
  ) 

# Render with gt
tbl1 <- table1_df |>
  gt() |>
  cols_label(
    trajectory = "PTSD Symptom Trajectory",
    k          = html("<em>k</em>"),
    N_pct      = "N (%)",
    bayes_prev = "Relative Prevalence (Upper Bound)",
    bayes_cri  = "95% CrI",
    cons_prev  = "Conservative Prevalence (Lower Bound)",
    cons_cri   = "95% CrI",
    logit      = "Logit Estimate",
    Tau        = html("<em>&tau;</em>"),
    Rhat_max   = "Max R-hat"
  ) |>
  cols_align("left",   columns = trajectory) |>
  cols_align("center", columns = c(k, N_pct, bayes_prev, bayes_cri, cons_prev, cons_cri, logit, Tau, Rhat_max)) |>
  fmt_integer(columns = k, use_seps = TRUE) |>
  fmt_number(columns = c(logit, Tau), decimals = 2) |>
  fmt_number(columns = Rhat_max, decimals = 3) |>
  opt_row_striping() |>
  tab_style(
    style = cell_text(size = px(11), font = "Times New Roman"),
    locations = cells_source_notes()
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_options(
    table.width = pct(100),
    data_row.padding = px(6),
    table.font.size = px(13),
    table.font.names = "Times New Roman"
  ) |>
  tab_source_note(
    source_note = html("<em>Note.</em> <em>k</em> represents the number of unique samples. Relative Prevalence (Upper Bound) reflects the posterior mean prevalence, logit estimate, and 95% credible interval (CrI) from a Bayesian random-intercept binomial GLMM (brms) with weakly informative priors (Intercept ~ Normal(0, 1.5); study SD ~ Student-t(3, 0, 1)), fit only to studies that identified the given trajectory. Conservative Prevalence (Lower Bound) fits the same model to all studies with a usable sample size, coding studies that did not identify the trajectory as contributing zero events. <em>&tau;</em> represents the between-study standard deviation. Max R-hat evaluates MCMC convergence, where values &le; 1.01 indicate good chain mixing and reliable posterior estimates.")
  ) |>
  opt_table_font(
    font = list(
      google_font("Times New Roman"),
      default_fonts()
    )
  )

tbl1

# Optional exports
export_dir <- "Prevalences Tables"
if (!dir.exists(export_dir)) {
  dir.create(export_dir, recursive = TRUE)
}

gt::gtsave(tbl1, file.path(export_dir,"prevalences-summary-table.png"))
gt::gtsave(tbl1, file.path(export_dir,"prevalences-summary-table.rtf"))
```

## Sensitivity Analyses Summary

```{r sensitivity-summary-table, message=FALSE, warning=FALSE}
mk_row_df <- function(r, label) {
    tibble(
      Analysis = label,
      k        = as.integer(r$Samples_included),
      `Relative Prevalence (95% CrI)` = 
        sprintf("%.1f%% [%.1f%% – %.1f%%]",
                100 * as.numeric(r$Pooled_Prevalence),
                100 * as.numeric(r$CrI_Lower),
                100 * as.numeric(r$CrI_Upper)),
      `Logit Estimate` = as.numeric(r$Logit_Estimate),
      Tau              = as.numeric(r$Tau)
    )
}

# Labels / mappings
trajectory_map <- c(
  "Decreasing"           = "Decreasing symptoms",
  "Partially_Decreasing" = "Partially decreasing symptoms",
  "Persistent"           = "Persistent symptoms",
  "Relapsing"            = "Relapsing symptoms"
)

analysis_map <- c(
  "Main"  = "Main analysis (all studies)",
  "sens1" = "Excluding studies with extreme prevalence (≥90%)",
  "sens2" = "Including only large-sample studies (N ≥ 1000)",
  "noinf" = "Excluding influential studies (Pareto k > 0.7)",
  "cons"  = "Conservative estimate (zero-imputed for non-identifying studies)"
)

# Build the spec of objects to look up
all_specs <- tidyr::crossing(
  trajectory_stub = names(trajectory_map),
  analysis_stub   = names(analysis_map)
) |>
  mutate(
    object_name = dplyr::case_when(
      analysis_stub == "Main"  ~ paste0("bglmm_", trajectory_stub),
      TRUE                     ~ paste0("bglmm_", trajectory_stub, "_", analysis_stub)
    ),
    Trajectory = trajectory_map[trajectory_stub],
    Analysis   = analysis_map[analysis_stub]
  )

# Build the table by reading the one-row data.frames created earlier
big_tbl <- purrr::pmap_dfr(
  all_specs,
  function(object_name, Trajectory, Analysis, ...) {
    obj <- tryCatch(get(object_name, envir = .GlobalEnv), error = \(e) NULL)
    if (!is.null(obj) && is.data.frame(obj) && nrow(obj) >= 1) {
      mk_row_df(obj, Analysis) |> mutate(Trajectory = Trajectory, .before = 1)
    }
  }
)

# Enforce display order
trajectory_order <- c(
  "Decreasing symptoms",
  "Partially decreasing symptoms",
  "Persistent symptoms",
  "Relapsing symptoms"
)

big_tbl_ordered <- big_tbl |>
  filter(!is.na(k) & k > 5) |>
  mutate(Trajectory = factor(Trajectory, levels = trajectory_order)) |>
  arrange(Trajectory)

# Create gt table
gt_all <- 
  gt(big_tbl_ordered, groupname_col = "Trajectory") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_row_groups()
  ) |>
  cols_label(
    Analysis = "Analysis",
    k = html("<em>k</em>"),
    `Relative Prevalence (95% CrI)` = "Relative Prevalence (95% CrI)",
    `Logit Estimate` = "Logit Estimate",
    Tau = html("<em>&tau;</em>")
  ) |>
  fmt_number(columns = c(`Logit Estimate`, Tau), decimals = 2) |>
  cols_align(align = "left", columns = Analysis) |>
  cols_align(align = "center", columns = c(k, `Relative Prevalence (95% CrI)`, `Logit Estimate`, Tau)) |>
  opt_row_striping() |>
  tab_options(
    table.font.names = c("Times New Roman", "Times", "serif"),
    data_row.padding = px(6),
    column_labels.font.weight = "bold",
    row_group.as_column = FALSE,
    table.font.size = px(13)
  ) |>
  tab_style(
    style = cell_text(size = px(11), font = "Times New Roman"),
    locations = list(cells_source_notes(), cells_footnotes())
  ) |>
  tab_source_note(md(
    "<em>Note.</em> <em>k</em> represents the number of unique samples. Logit estimate reflects pooled prevalence on a log-odds scale. <em>&tau;</em> indicates between-study standard deviation. All estimates rely on Bayesian GLMMs (brms) with weakly informative priors. The conservative (zero-imputed) analysis codes studies that did not identify a given trajectory as contributing zero events, rather than excluding them, and can be read as a lower-bound estimate against the main analysis' upper bound. Analyses based on 5 or fewer studies (<em>k</em> &le; 5) are omitted from this table, as random-effects estimates become unstable with very few clusters."
  ))

# Footnotes for excluded influential studies 
.influential_for <- function(traj_stub) {
  nm <- paste0("influential_studies_", traj_stub)
  v  <- tryCatch(get(nm, envir = .GlobalEnv), error = \(e) NULL)
  if (is.null(v)) character(0) else as.character(v)
}
format_note <- function(vec) {
  if (length(vec) == 0) "No influential studies excluded."
  else paste0("Excluded influential studies: ", paste(vec, collapse = "; "))
}

excluded_by_group <- setNames(
  lapply(names(trajectory_map), .influential_for),
  nm = unname(trajectory_map[names(trajectory_map)])
)
excluded_notes <- lapply(excluded_by_group, format_note)

# add notes to table
gt_all <- purrr::reduce(
  names(excluded_notes),
  .init = gt_all,
  .f = function(gt_obj, grp_label) {
    tab_footnote(
      gt_obj,
      footnote  = excluded_notes[[grp_label]],
      locations = cells_body(
        columns = "Analysis",
        rows = (Trajectory == grp_label) & (Analysis == "Excluding influential studies (Pareto k > 0.7)")
      )
    )
  }
)

gt_all

# Exports
gt::gtsave(gt_all, file.path(export_dir, "prevalences-sensitivity-table.rtf"))
gt::gtsave(gt_all, file.path(export_dir,"prevalences-sensitivity-table.png"), vwidth = 700) 
```
